---
title: "615 Midterm Project"
author: "Jingrong Cheng"
date: "10/21/2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyr)
library(knitr)
library(arm)
library(foreign)
library(reader)
library(ggplot2)
library(latticeExtra)
library(lubridate)
library(data.table)
library(animation)
library(ggthemes)
```

This data cleaning & EDA project was completed by Jingrong Cheng and Tianwen (Tina) Huan.
We are curious about how data reflects about climate change, so we found a suitable dataset on Kaggle.com.
https://www.kaggle.com/berkeleyearth/climate-change-earth-surface-temperature-data
We selected "Global Land Temperatures By City" file to do this project. 

```{r}
CT <- read.csv("~/Desktop/GlobalLandTemperaturesByCity.csv")
```

## 1. Check the original dataset
First of all, we want to know the basic information of this file, so we apply "summary","dim","str" function.
```{r}
summary(CT)
dim(CT)
str(CT$Country)
```

Details highlighted:
*year/month goes form 1743.11 to 2013.09 
*8599211 rows, 159 Countries 
Due to the dataset is too large, we decided to select a subset data only focusing on United States and from year 1900.

## 2. Choose the subset
```{r}
cityT <- CT %>%
     filter(Country=="United States") %>% # narrow down to United States
     mutate(date=dt) %>%
     separate(dt, c("year", "month", "day")) %>% # sepearate the year month and day
     filter(year >= 1900)  # select the data after year 1990
  
# drop the "day" and "Country" columns
cityT <- subset(cityT, select = c(10,1,2,4,5,6,8,9))

# check missing data
cityT[!complete.cases(cityT),]

# drop all the data for 2013.09
cityT <- cityT %>%
         filter(year!="2013" | month!="09")
```

The initial cleaned dataset includes 350805 observations and 9 variables.
All the columns: X, date, year, month, Average temperature, average temperature uncertainty, city, latitude, longitude.
The order of date is from 1900/1/1 to 2013/8/1.

## 3. Data character transformation
In order to explore data eaiser, we transformed the class of some columns.
```{r}
cityT$date<-as.Date(cityT$date,"%Y-%m-%d")
cityT$year<-as.numeric(cityT$year,"%Y")
cityT$month<-as.numeric(cityT$month,"%m")


cityT$lat<-as.numeric(gsub("N|E|S|W", "",cityT$Latitude))*ifelse(grepl("S",cityT$Latitude),-1,1)
cityT$long<-as.numeric(gsub("N|E|S|W", "", cityT$Longitude))*ifelse(grepl("W",cityT$Longitude),-1,1)

cityT <- data.table(cityT)

# remove Hawaii & Alaska
cityT <- cityT %>%
         filter(long>=-130 & lat>=25 & lat<=55)
```

## 4. Location Graph According to latitudes and longitude
Using groups of latitudes and longitudes to practice "ggplot" function.
```{r}
citylocation <- subset(cityT, select = c("City","lat","long"))
citylocation <- citylocation %>% distinct(.keep_all= FALSE)


ggplot(citylocation, aes(long, lat), col=temp) +
  borders("state") + geom_point()+
  scale_size_area() + coord_quickmap() +
  labs(x="Longitude", y="Latitude", title="Location of Cities")
```


## 5. US Temperature by Month and Year 1900-2012
```{r}
# choose subset
aT <- cityT %>%
       group_by(year, month) %>%
       summarise(temp=mean(AverageTemperature))

# month trend for different year
ggplot(aT, aes(x=month, y=temp)) +
  geom_jitter(aes(colour=year)) + ggtitle("US Temperature by Month and Year") +
  labs(x="Month", y="Average Temperature")

# year trend for different month
ggplot(aT, aes(x=year, y=temp)) +
  geom_jitter(aes(colour=factor(month))) + 
  ggtitle("US Temperature by Month and Year") +
  labs(x="Month", y="Average Temperature")
```


## 6. Boston Temperature by Month and Year 1900-2012
```{r}
# choose subset
aTB <- cityT %>%
       filter(City=="Boston") %>%
       group_by(year, month)

# month trend for different year
ggplot(aTB, aes(x=month, y=AverageTemperature)) +
  geom_jitter(aes(colour=year)) + 
  ggtitle("Boston Temperature by Month and Year") +
  labs(x="Month", y="Average Temperature")

# year trend for different month
ggplot(aTB, aes(x=year, y=AverageTemperature)) +
  geom_jitter(aes(colour=factor(month))) + 
  ggtitle("Boston Temperature by Month and Year") +
  labs(x="Month", y="Average Temperature")

# year trend for January
BostonT<- cityT %>% filter(City == "Boston")
bostonjan<-BostonT %>% filter(month == 1)
ggplot(bostonjan, aes(x=year, y=AverageTemperature))+
  geom_line() +
  labs(x="Month", y="Average Temperature", title="Boston January Temperature")
```

## 7. US Temperature by Year 1900-2012
```{r}
# choose subset
aTy <- cityT %>%
       group_by(year) %>%
       summarise(temp=mean(AverageTemperature))

# year trend
ggplot(aTy, aes(x=year, y=temp)) +
  geom_jitter(aes(colour=year)) + 
  ggtitle("US Temperature by Year") + geom_smooth() +
  labs(x="Year", y="Average Temperature")
```

## 8. Horizonplot
```{r}
y10 <- subset(cityT, year %in% c(1900,1910,1920,1930,1940,1950,1960,1970,1980,1990,2000,2010))

citytable1 <- y10 %>%
             filter(City=="Boston" | City=="Chicago" | City=="Los Angeles" | City=="Madison") %>%
             group_by(year, month, City) %>%
             summarise(temp=AverageTemperature) %>%
             spread(City, temp) 

citytable <- subset(citytable1, select=c(3:6))
             
horizonplot(ts(citytable), horizonscale = 15, colorkey = TRUE,
            xlab="1900,1910,1920,1930,1940,1950,1960,1970,1980,1990,2000,2010")
```

## 9. Heatmap for 5 main Cities
```{r}
# choose the subset
hm <- cityT %>% 
        group_by(month, City) %>%
        summarise(temp=mean(AverageTemperature)) %>%
        filter(City=="Boston" | City=="Chicago" | 
                 City=="Los Angeles" | City=="New York" | City=="Seattle")

# Heatmap
ggplot(hm, aes(x=month, y=City, fill=temp, frame=City)) +
  geom_tile(color="white", size=0.1) +
  scale_fill_gradient(name="Average Temperature", low="white", high="red") +
  coord_equal() +
  labs(x = "Months", y = "", title = "Average Temp from 1900 - 2012") +
  theme_tufte() +
  theme(axis.ticks = element_blank()) +
  theme(axis.text = element_text(size = 14)) +
  theme(plot.title = element_text(size = 15)) +
  theme(legend.title = element_text(size = 15)) +
  theme(legend.text = element_text(size = 10))
```

## 10. Temperature Trend by Season
```{r}
s <- function(m){
            factor((m %% 12) %/% 3, labels = c('Winter', 'Spring', 'Summer', 'Autumn'))}

par(mfrow=c(1,2))
# US Tempreture by Season
sus <- cityT %>%
       group_by(year, season=s(month)) %>%
       summarise(temp=mean(AverageTemperature))

sus %>%
    ggplot(mapping=aes(x=year, y=temp)) +
     geom_line(mapping = aes(color=season)) +
     theme_bw() + labs(x="Year", y="Average Temperature", 
                       title="US Season Temperature 1900-2012")

# Boston Tempreture by Season
sbo <- cityT %>%
       filter(City=="Boston") %>%
       group_by(year, season=s(month)) %>%
       summarise(temp=mean(AverageTemperature))

sbo %>%
    ggplot(mapping=aes(x=year, y=temp)) +
     geom_line(mapping = aes(color=season)) +
     theme_bw() + labs(x="Year", y="Average Temperature", 
                       title="Boston Season Temperature 1900-2012")
```

